Data Manipulation Language for fuzzy temporal databases.

### 2013-02-15 joins/time.tex
update!

\subsubsection{Representation}
Several proposals for managing uncertain time in a database exist. Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Garrido2009}, \cite{Galindo2001}. In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004} and \cite{Nagypal2003}. In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees. Also, in order to work properly with fuzzy operators, the underlying domain should be numeric. In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Husfeld1996}.

If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval. Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig. \ref{fig:fuzzy-validity-period}):
\begin{itemize}
\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
\end{itemize}

>>>>

\subsubsection{Representation}
Several proposals for representing uncertain time in a database exist. Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Garrido2009}, \cite{Galindo2001}. In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004} and \cite{Nagypal2003}. In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees. Also, in order to work properly with fuzzy operators, the underlying domain should be numeric. In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Husfeld1996}.

If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
\begin{definition}
A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval. Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig. \ref{fig:fuzzy-validity-period}):
\begin{itemize}
\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
\end{itemize}
\end{definition}

\begin{definition}
A \emph{Possibilistic Valid-time Period} is an ill-known interval in time, which models a time period during which an object in a certain state is valid.
\end{definition}

Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start and/or end point of a time interval (and thus about the time interval itself) if such uncertainty exists. The interpretation is \emph{disjunctive}: the PVP represents exactly one valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown. In the presented model, only PVPs are considered of which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have exactly the same characteristics as the membership functions of fuzzy numbers. A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists ! x : \mu_{P}(x) > 0$.

As mentioned in \cite{Pons2011}, this approach differs from the one where a valid-time period is represented by one fuzzy set. Such a fuzzy set is seen as a possibility distribution on $\mathbb{R}$ and thus defines just one ill-known value. However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\Pow(\mathbb{R})$.

### 2013-02-15 temporaldml/main.tex
update!

\item We will note by $r$ any valid instance of $R$. 
      \begin{equation}
       \label{eq:valid-time-instance}
      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ T x\ T
      \end{equation}

@@

\section{\label{sec:selection}Selection}
The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints from a given relation $R$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
\begin{equation}
 \label{eq:selection}
\sigma_{P} \left( R \right)
\end{equation}

Where $R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
% \begin{itemize}
% \item relation.attribute boolean relation relation.attribute
% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
% \end{itemize}
\begin{equation}
 \label{eq:selection_formula}
P = \left \lbrace Q, Q^{t}\right \rbrace
\end{equation}

Where $q$ is a boolean combination of (possibly) fuzzy (and bipolar) non-temporal constraints and $q^{t}$ is also a boolean combination of fuzzy temporal constraints.

\begin{equation}
 \label{eq:non-temporal-constraints}
Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right \rbrace
\end{equation}
Where $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$ and val is a value in the domain of the attribute.

>>>>

\item We will note by $r$ any valid instance of $R$. 
      \begin{equation}
       \label{eq:valid-time-instance}
      r \subseteq D_1\times \ \ldots\ \times \ D_n \times \ T \times \ T
      \end{equation}

@@

\section{\label{sec:selection}Selection}
The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints $P$ from an instance $r$ of the relation $R \subseteq
 A \times I$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
\begin{equation}
 \label{eq:selection}
\sigma_{P} \left( r \right)
\end{equation}

Where $r \in R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
% \begin{itemize}
% \item relation.attribute boolean relation relation.attribute
% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
% \end{itemize}
\begin{equation}
 \label{eq:selection_formula}
P = \left \lbrace Q, Q^{T}\right \rbrace
\end{equation}

Where $Q$ is the set of (possibly) fuzzy (and bipolar) non-temporal constraints and $Q^{T}$ is also a boolean combination of fuzzy temporal constraints.

\begin{equation}
 \label{eq:non-temporal-constraints}
Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{a}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{a}_n}  \theta \mbox{ val } \right \rbrace
\end{equation}

Where:
\begin{itemize}
 \item $q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{a \in A }$ is an atomic constraint. A possitive preference is notated by $q^{\left \lbrace \mbox{pos} \right \rbrace}$ whether a negative preference is notated by $q^{\left \lbrace \mbox{neg} \right \rbrace}$. The constraint refers to an attribute $a$ that belongs to the set of attributes $A \subseteq A_1 \ldots A_n $ of the relation $R$.
\item $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$.
\item \emph{val} is a value in the domain of the queried attribute. 
 \end{itemize}
